---
title: "Krill 2019 Respirometry"
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: cerulean
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '4'
editor_options:
  chunk_output_type: console
---
# Load libraries
```{r libraries}
library(tidyverse)
library(here)
library(janitor)
```
# Read and clean data
```{r read-and-clean-data}
#vial_volume in liters
vial_vol <- 0.02806

d_resp <- read_csv(here("data_raw", "KRILL_Resp_alltrials.csv")) %>%
  clean_names() %>%
  select(date, time, sensor_name, delta_t, value, temp) %>%
  rename(krill_id = sensor_name,
         oxygen = value) %>%
  mutate(oxygen = as.numeric(oxygen)) %>%
  #Convert from mg/L to ug/vial
  mutate(oxygen = oxygen  * 0.02806 * 1000) %>%
  # convert delta_t from minutes to hours
  mutate(delta_t = delta_t / 60) %>%
  #remove some crazy values at the end of one run
  filter(oxygen < 1000)
```
#Correct for mis-entered temperature data
#TODO fix this

# Plot oxygen curves
```{r oxygen-curves}
d_resp %>%
  ggplot(aes(delta_t, oxygen)) +
    geom_point(aes(color = krill_id)) 
```
# Slope without outliers function
```{r slope-no-outlier-fun}
slope_no_outlier_fun <- function(x, y){
  d <- data.frame(x,y)
  m <- lm(y~x, data = d)
  sample_size <- nrow(d)
  d_filter <- d %>%
    mutate(cooksd = cooks.distance(m)) %>%
    mutate(outlier = if_else(cooksd > 4 /sample_size, TRUE, FALSE)) %>%
    filter(!outlier)
  m_filter <- lm(y~x, data = d_filter)
  return(m_filter$coefficients[2])
}
```

# Calcluate slopes
```{r calc-slopes}

#hours of acclimation 
acclim_t <- 10/60
#duration of exposure (hours)
duration <- 1.5

d_slopes <- d_resp %>%
  filter(delta_t > acclim_t) %>%
  filter(delta_t < duration) %>%
  group_by(krill_id) %>%
  summarise(oxy_slope = slope_no_outlier_fun(delta_t, oxygen))

```
# Plot oxygen slopes by krill
```{r oxy-slope-plot}
d_slopes %>%
  ggplot(aes(krill_id, oxy_slope)) +
    geom_point()
```
# Read krill info
```{r read-krill-info}
d_krill <- read_csv(here("data_raw", "RespirometryTrials_all.Animal.Info.csv")) %>%
  clean_names() %>%
  rename(krill_id = sensor_name)
```
# Merge slopes and krill info
```{r merge-slopes-and-info}
d_slope_krill <- d_krill %>%
  full_join(d_slopes, by = "krill_id")
```
# Get blanks data frame
Average slope over the two blanks per trial
```{r blanks-data}
d_blanks <- d_slope_krill %>%
  filter(moats == "blank") %>%
  group_by(trial_id) %>%
  summarise(blank_oxy_slope = mean(oxy_slope))
```
# Complete data frame
Assign blanks to krill trial do blank corrections
Standardize by size
```{r complete-data}
d <- d_slope_krill %>%
  left_join(d_blanks, by = "trial_id") %>%
  filter(moats != "blank") %>%
  #subtract the blanks
  mutate(oxy_slope_blank_adj = oxy_slope - blank_oxy_slope) %>%
  mutate(size = as.numeric(size)) %>%
  mutate(oxy_slope_blank_size_adj = oxy_slope_blank_adj / size) %>%
  # units of oxy_slope_blank_size_adj are mg of O2 per minute removed from vial per gram of krill
  # conver to ug of O2 consummed per hour per gram of krill
  mutate(repsiration = -(oxy_slope_blank_size_adj)) %>%
  #remove ambient treatment
  filter(treatment != "AMB")
```

# Poster plot
```{r poster-plot}
# oceans in high so2 poster plot

d %>%
  filter(repsiration > 0 ) %>%
  mutate(Treatment = factor(treatment, levels = c("CUR", "TMP", "CHG"))) %>%
  ggplot(aes(Treatment, repsiration, color = Treatment)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) + 
  geom_jitter(shape=16, position=position_jitter(width = 0.2, height = 0), size = 2) +
  scale_color_manual(values = c("mediumblue", "orange1", "limegreen")) +
  scale_x_discrete(labels = c("Current Ocean", "Heat Wave", "Future Ocean")) +
  xlab("Treatment") + 
  ylab(expression(Respiration~(µg~O[2]~h^-1~g^-1))) +
  theme_bw(base_size = 32) +
  theme(legend.position="none")

ggsave(here("figs", "krill_respiration_poster.jpeg"))

```


